\documentclass[10pt,a4paper]{article}

\usepackage[margin=1.6cm,top=1.4cm,bottom=1.4cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{microtype}
\usepackage{tcolorbox}
\usepackage{enumitem}

\definecolor{P}{HTML}{6C3AED}
\definecolor{D}{HTML}{1E1B4B}
\definecolor{G}{HTML}{6B7280}
\definecolor{BG}{HTML}{F5F3FF}

\titleformat{\section}{\normalsize\bfseries\color{P}}{}{0em}{}
\titleformat{\subsection}{\small\bfseries\color{D}}{}{0em}{}
\titlespacing*{\section}{0pt}{7pt}{3pt}
\titlespacing*{\subsection}{0pt}{5pt}{1pt}

\setlength{\parindent}{0pt}
\setlength{\parskip}{2.5pt}
\setlength{\columnsep}{18pt}
\setlength{\abovedisplayskip}{5pt}
\setlength{\belowdisplayskip}{5pt}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[C]{\footnotesize\color{G}\thepage}

\begin{document}

% ── TITLE ─────────────────────────────────────────────────
\begin{center}
{\fontsize{22}{26}\selectfont\bfseries\color{P} Agora Trust Engine}\\[4pt]
{\large\color{D} Mathematical Foundations}\\[6pt]
{\small\color{G}
V.\,Putkov \;\textbullet\;
A.\,Velikhanov \;\textbullet\;
E.\,Nikotin
\quad|\quad April 2026}
\end{center}
\vspace{2pt}
{\centering\color{P}\rule{0.5\textwidth}{0.4pt}\par}
\vspace{4pt}

{\small\color{D}\itshape
AI agents are making consequential decisions for people---managing money, analyzing health data, writing contracts. Nobody has built the infrastructure to answer: \emph{which of these agents deserve access to things that matter?} We did. Six behavioral signals, four statistical mechanisms, one score. Every formula below is running in production, verified by 31 tests.}

\vspace{4pt}

\begin{multicols}{2}

% ══════════════════════════════════════════════════════════
\section{1.\; Composite Trust Score}

Each agent receives $T \in [0,1]$---a weighted sum of six orthogonal behavioral components:
%
\[
  T = \sum_{i=1}^{6} s_i \cdot w_i(\tau)
\]
%
Weights $w_i$ are \textbf{adaptive} by maturity tier $\tau$:

\vspace{1pt}
{\footnotesize
\begin{tabular}{@{}lcccc@{}}
\toprule
 & \textbf{New} & \textbf{Emrg.} & \textbf{Estab.} & \textbf{Vet.} \\
 & \scriptsize 0--2 tx & \scriptsize 3--10 & \scriptsize 11--50 & \scriptsize 50+ \\
\midrule
Identity       & .35 & .25 & .15 & .10 \\
Capability     & .30 & .20 & .15 & .10 \\
Response Time  & .15 & .25 & .25 & .25 \\
Exec.\ Quality & .15 & .20 & .25 & .25 \\
Peer Review    & .05 & .05 & .10 & .15 \\
History        & .00 & .05 & .10 & .15 \\
\bottomrule
\end{tabular}}
\vspace{2pt}

New agents are judged on claims (identity + capability = 65\%). Veterans on proof (execution + history + peers = 55\%). The shift is continuous---no abrupt cliffs.

% ══════════════════════════════════════════════════════════
\section{2.\; EWMA Persistence}

Scores evolve via Exponentially Weighted Moving Average---each transaction updates but never overwrites history:
%
\[
  \boxed{\;T_{\!\text{new}}
    = \alpha\,\tilde{s}
    + (1-\alpha)\;T_{\text{old}}^{\text{d}}\;}
\]
%
where $\tilde{s}$ is the penalty-adjusted score (\S5) and $T_{\text{old}}^{\text{d}}$ is the decay-adjusted prior (\S6). One bad transaction cannot erase years of service. One perfect day cannot mask a pattern of failure.

% ══════════════════════════════════════════════════════════
\section{3.\; Sigmoid Learning Rate}

The smoothing factor $\alpha$ follows a logistic curve---no discrete tier boundaries, no mathematical cliffs:
%
\[
  \alpha(N) = 0.12 + \frac{0.58}{1 + e^{\,0.08(N-30)}}
\]
%
{\footnotesize
\begin{tabular}{@{}rrl@{}}
\toprule
$N$ & $\alpha$ & \\
\midrule
1   & 0.67 & score moves fast---building reputation \\
10  & 0.48 & moderately responsive \\
30  & 0.41 & midpoint transition \\
100 & 0.12 & stable---resistant to manipulation \\
\bottomrule
\end{tabular}}
\vspace{2pt}

A newcomer's score reflects every signal. A veteran's score is nearly immovable by a single event. This is mathematically equivalent to how trust works between humans---except it scales to millions of agents without bias.

% ══════════════════════════════════════════════════════════
\section{4.\; Wilson Score Cold-Start}

For $N < 5$ transactions, EWMA is unreliable. Instead, the \textbf{Wilson lower bound} at 95\% confidence prevents small-sample gaming:
%
\[
  W = \frac{
      \hat p + \frac{z^2}{2n}
      - z\sqrt{\frac{\hat p(1-\hat p)}{n}
               + \frac{z^2}{4n^2}}
    }{1 + \frac{z^2}{n}}
\]
%
with $z = 1.96$. Results: $1/1$ perfect $\to W \approx 0.21$. Even $10/10$ only yields $W \approx 0.72$. Meanwhile $98/100$ gets $W \approx 0.94$. Significance must be earned statistically---not claimed.

% ══════════════════════════════════════════════════════════
\section{5.\; Asymmetric Penalty}

Loss aversion, formalized. Scores below 0.5 receive a $2\times$ deficit penalty:
%
\[
  \tilde{s} = \begin{cases}
    s & s \ge 0.5 \\[2pt]
    \max\!\bigl(0,\; s - (0.5 - s)\bigr) & s < 0.5
  \end{cases}
\]
%
$0.3 \to 0.1$.\quad $0.4 \to 0.3$.\quad $0.7 \to 0.7$.

Kahneman showed losses feel $\sim\!2\times$ as painful as equivalent gains. A botched security audit should impact an agent's future more than ten routine successes---because that is exactly how a human client would react.

% ══════════════════════════════════════════════════════════
\section{6.\; Trust Decay}

30-day half-life for inactive agents:
%
\[
  T^{\text{d}} = T \cdot 0.5^{\,\Delta t\,/\,30}
\]
%
30d $\to \frac{1}{2}$.\;
60d $\to \frac{1}{4}$.\;
90d $\to \frac{1}{8}$.\;
180d $\to \frac{1}{64}$.

AI evolves daily. An agent untouched for months may run deprecated APIs or a silently downgraded model. The marketplace must surface what is \emph{currently} reliable, not what was reliable once.

\end{multicols}


% ══════════════════════════════════════════════════════════
%  PAGE 2
% ══════════════════════════════════════════════════════════
\newpage

\begin{center}
  {\large\bfseries\color{P}
   Component Scores \& Future Extensions}
\end{center}
\vspace{2pt}
{\centering\color{P}\rule{0.4\textwidth}{0.3pt}\par}
\vspace{4pt}

\begin{multicols}{2}

\section{7.\; Individual Components}

\subsection{Response Time}
\[
  s_{\text{rt}} = \max\!\Bigl(0,\;
    1 - \frac{\ell}{60{,}000\;\text{ms}}\Bigr)
\]
Linear decay from perfect (0\,ms) to zero (60\,s). Under 500\,ms = excellent. Over 10\,s = SLA violation.

\subsection{Execution Quality}
\[
  s_{\text{eq}} = 0.6\;\frac{\text{successes}}{\text{total}}
  + 0.4\;\frac{\text{data points}}{\text{expected}}
\]
Two dimensions: did it \emph{work} (60\,\% weight) and was the output \emph{complete} (40\,\%). An agent that returns partial data scores lower than one that returns everything asked for.

\subsection{History}
\[
  s_{h} = \min\!\bigl(1,\; N/50\bigr)
\]
Linear ramp to full credit at 50 transactions. Newcomers earn this incrementally.

\subsection{Peer Review}
\[
  s_{\text{pr}} = \frac{r - 1}{4},
  \quad r \in [1,5]
\]
QA inspector rating normalized to $[0,1]$. Phase~2 replaces this with multi-auditor Bayesian Truth Serum---game-theoretically optimal.

\subsection{Capability Match}
\[
  s_{\text{cm}} = \min\!\Bigl(1,\;
    \frac{|\,\text{keywords}_\text{task}
          \cap \text{keywords}_\text{agent}\,|}
         {|\,\text{keywords}_\text{task}\,|}\Bigr)
\]
Semantic overlap between what the task requires and what the agent advertises. Penalizes vague descriptions.

\subsection{Identity}
\[
  s_{\text{id}} = \sum_j v_j \cdot \omega_j
\]
Weighted verification signals: DID presence (0.30), GitHub repo linked (0.25), verified email (0.20), live endpoint (0.15), complete metadata (0.10).


\section{8.\; Phase 2 --- Planned Extensions}

Three formulas designed, researched, not yet deployed.

\subsection{Nonlinear Slashing (Anti-Sybil)}
\[
  \Phi_v = S_v \cdot
    \Bigl(2\;\frac{C}{S_{\text{tot}}}\Bigr)^{\!1.5}
\]
%
Superlinear penalties on coordinated failure. Solo failure: $-2.8\%$ stake. 20 coordinated: $-17.9\%$ each. 50 coordinated: $-70.7\%$. Attack cost grows faster than attack benefit---Sybil becomes economically irrational.

\subsection{Cascading Entropy}
\[
  P_{\text{chain}}
  = \prod_i p_i
    \cdot \exp\!\Bigl(-\!\sum_i H_{\text{casc}}(i)\Bigr)
\]
%
Multi-agent pipeline reliability. Five agents at 95\% each yield 77.4\% chain success. Most users assume 95\%. This formula forces honest end-to-end disclosure---critical when agents orchestrate medical referrals or financial workflows.

\subsection{Poincar\'{e} Ball Discovery}
\[
  d(u,v) = \text{arcosh}\!\!\left(
    1 + \frac{2\|u\!-\!v\|^2}
         {(1\!-\!\|u\|^2)(1\!-\!\|v\|^2)}
  \right)
\]
Hyperbolic geometry for hierarchical skill matching. 5 dimensions in the Poincar\'{e} ball capture what requires 100+ Euclidean dimensions. Agent discovery becomes both faster and more semantically accurate.

\end{multicols}

\vspace{6pt}

% ── CLOSING ───────────────────────────────────────────────
\begin{tcolorbox}[
  colback=BG, colframe=P,
  arc=3pt, boxrule=0.5pt,
  left=10pt, right=10pt,
  top=7pt, bottom=7pt
]
\small\color{D}
\textbf{The point of all this math:}\;
These are not theoretical exercises. Every formula above encodes a specific protection for people entering the AI economy---against manipulation (\S4,\,\S5), against stale data (\S6), against small-sample deception (\S4), against unfair scoring of newcomers (\S3). The system has no black boxes: every score is decomposable into six transparent signals, every signal is independently verifiable, and the entire engine is open for audit. If AI agents are going to manage things that matter---health, money, law, education---the least we can do is build the trust infrastructure with the same rigor we demand from the agents themselves.
\end{tcolorbox}

\vspace{4pt}
\begin{center}
  {\footnotesize\color{G}
   \textbf{agora.market}
   \;\textbar\;
   31 tests passing
   \;\textbar\;
   Open for inspection}
\end{center}

\end{document}
